---
title: 'Chapter 13: Simple Regression'
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## 

```{r}

# set current directory, then read data file

realEstate = read.csv("RealEstate.csv")
colnames(realEstate) # very long variable names
head(realEstate)

realEstate <- realEstate |> dplyr::select(-No) #remove first column
head(realEstate)

y = realEstate$Y.house.price.of.unit.area
x1 = realEstate$X1.transaction.date
x2 = realEstate$X2.house.age
x3 = realEstate$X3.distance.to.the.nearest.MRT.station
x4 = realEstate$X4.number.of.convenience.stores
x5 = realEstate$X5.latitude
x6 = realEstate$X6.longitude

#or

namevec <- c("x1", "x2", "x3", "x4", "x5", "x6", "y")
names(realEstate) <- namevec
realEstate



```

## Exploring dataset

```{r}

#install.packages("GGally")    # only run once
library(GGally)

ggpairs(realEstate)

ggpairs(realEstate |> select('x3', 'y'))

#or ggpairs(realEstate[, c(3, 7)])



```

Plot y and x3 and try to observe the form of the relationship 

```{r}

plot(x3,y)
#plot(y,x3)

```

Plot log(y) and x3 to see if the relationship becomes more linear:

```{r}

plot(x3,log(y))

```
To see that the form between x3 and y is less linear than the form between x3 and log(y), use the correlation coefficient r:

```{r}

cor(x3,y)

cor(x3,log(y)) # close to -1 than cor(x3, y)

```

Find the linear regression line for log(y) and x3:

```{r}

x3lm = lm(y ~ x3)

#x3lm
lm(y ~ x3)

## Assess the validity of the model, use the summary function:

#options(scipen=999)  # not show scientific notation

summary(x3lm)

```
Residual Analysis

```{r}

#get list of residuals 
res <- resid(x3lm)
#res
qqnorm(res)
qqline(res)

shapiro.test(res)

ggplot(data = NULL, aes(x = res)) + geom_density()


#produce residual vs. fitted plot
plot(fitted(x3lm), res)

#add a horizontal line at 0 
abline(0,0)


#install.packages("MASS")
library(MASS) #if it's not installed, do so.

#calculate studentized residuals, which is a scaled version of the residuals
stud_res <- studres(x3lm)
#stud_res
#res

#produce residual vs. fitted plot
plot(fitted(x3lm), stud_res,  ylab='Studentized Residuals', xlab='predicted values')

#add a horizontal line at 0 
abline(h = c(-3,0,3), lty = 2)


```


## Plot y and x5

```{r}

plot(x5,y)
cor(y,x5)

```

## Plot log(y) and x5

```{r}

plot(x5,log(y))
cor(log(y), x5)

```

## Compare the linear correlation between x5 and y and log(y)

```{r}

cor(x5,y)

cor(x5,log(y))

```

## Create a regression model for log(y) and x5

```{r}

x5model <- lm(y~x5)

x5model
## Check to see if the model is adequate. 

summary(x5model)


```

## How to use r-squared

```{r}

## change in y

45 - 35

## How much of this change can be explained by the model: Multiple R-squared is 
## 0.2985

10*.2985

#Response prediction
#define new observation
newdata = data.frame(x5=24.95)

#use model to predict a response value with the prediction interval
predict(x5model, newdata, interval = "predict", level = 0.95)
#predict(x5model, newdata, interval = "predict")

predict(x5model, newdata, interval = "predict", level = 0.90)


# For multiple values of x5's

newdata = data.frame(x5=c(24.95, 25.5, 30))

#use model to predict a response value with the prediction interval
predict(x5model, newdata, interval = "predict")

#This is which the band on the ggplot came from.


```

## A different "prediction" interval

```{r}

#use model to predict the mean response value with the confidence interval
predict(x5model, newdata, interval = "confidence")

```

## Another example

```{r}

ages = c(2, 2.5, 4, 5, 9, 11, 13, 15, 16, 18)
heights = c(2.3, 3, 3.5, 3.5, 4, 4.6, 5.1, 5.2, 5.6, 5.5)

plot(ages, heights)
cor(ages, heights)
cor(heights, ages)

lm(heights~ages)

plot(ages, heights, xlim = c(0, 20))
abline(lm(heights~ages))

## or
modelA <- lm(heights~ages)
modelA

```


## Linear regression, but the equation has a quadratic term

```{r}

xvals <- c(2, 4, 6, 8, 10, 12, 14)
yvals<- c(5,10,15,22,17,12,3)

plot(xvals,yvals)

#Simple Regression

model1 <- lm(yvals ~ xvals)

summary(model1)

xsq <- xvals^2


#Multiple Regression

model2 <- lm(yvals ~ xvals + xsq)
#model2

summary(model2)


```


```{r}

```

