---
title: "Bootstrapping in R"
output: html_document
#output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
getwd()

```

```{r bootstrap}

data_to_bootstrap <- read.csv("bootstrap_sample.csv")
head(data_to_bootstrap)
#dim(data_to_bootstrap)

# Save the variable column in the data set to a vector named original_sample
original_sample = data_to_bootstrap$Variable

# Create an empty vector to store all of the sample standard deviations:
boot_sample = numeric()

# or boot_sample <- NULL

# Determine the length of the sample and call this value n
n <- length(original_sample)
n

# Write a for-loop which computes the standard deviation on 10000 samples 
# of size n taken/bootstrapped from the original sample.


for(i in 1:10000){
  boot_sample[i]= sd(sample(original_sample, n, replace = TRUE))
}


# To estimate the sample standard deviation, you take an average of all of 
# the sample standard deviations

mean(boot_sample)


# Find a 90% confidence interval for the standard deviation

# Use the formula low <- (1 - confidence)/2 and high <- 1 - low

confidence <- .90
low <- (1 - confidence)/2
high <- 1 - low
c(low, high)

quantile(boot_sample,c(low,high))

# Define the lower bound of the confidence interval 
lower = quantile(boot_sample,low)
upper = quantile(boot_sample,high)

# Plot a histogram of the sampling distribution and insert vertical lines
# at the confidence interval bounds. 
hist(boot_sample,main="Bootstrapped Sample Dist. for Std. Dev",
     probability = TRUE)
abline(v=lower,col="red")
abline(v=upper,col="red")

# For practice, you should try to reproduce the histogram with ggplot2.

```

```{r }
```

Another Example: Bootstrapped Confidence Interval for population mean $\mu$

Consider again the bootstrap_sample.csv data that we have read into R.

1. Compute the sample mean.

```{r}
mean(original_sample)
```

2. Create an empty vector called boot_mean which will contain 10000 bootstrapped sample means.

```{r}
# Empty vector which will contain all of the bootstrapped sample means
boot_mean = numeric()
```

3. Write a bootstrap which computes 10000 sample means and stores the observed values of the statistic in the vector boot_mean.

```{r}
for(k in 1:10000){
  temp_samp = sample(original_sample,n,replace=TRUE)
  temp_mean = mean(temp_samp)
  boot_mean[k] = temp_mean
}
```

4. Use the bootstrapped sampling distributed to compute a 95% bootstrapped confidence interval
for $\mu$.

```{r}
quantile(boot_mean,c(0.025,0.975))
```

5. Compare the mean of the single sample in bootstrap_sample.csv and the mean of the 10000 bootstrapped sample. What do you observe?

```{r}
mean(boot_mean)

mean(original_sample)
```

6. Compare the variance of the single sample in bootstrap_sample.csv and the variance of the 10000 bootstrapped sample. What do you observe?

```{r}
var(boot_mean)

var(original_sample)

sig.os <- var(original_sample)*(n-1)/n

sig.os/var(boot_mean)
```

7. Plot the histogram of the 10000 bootstrapped sample means. What do you observe?

```{r}
ggplot(data = NULL, aes(x = boot_mean)) + geom_histogram(binwidth = 2)
```

```{r}

```

