---
title: "Normal Distribution"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## demo of a histogram

```{r}

# always do this.
library(tidyverse)

```


```{r}

nd <- rnorm(500, mean=50, sd=5)
nd

# when not working with data frames

ggplot(NULL, aes(x = nd)) + geom_histogram(binwidth = 3, color="black", fill="gray") + ggtitle("Distribution of X") + labs(x = "Values of X")

```

## Basic R histogram and density curve.

```{r}
CO2

Y=CO2$uptake

hist(Y, main="CO2 uptake", xlab = "Uptake Rate", prob = FALSE, col = "yellow")

hist(Y, main="CO2 uptake", xlab = "Uptake Rate", prob = TRUE, col = "yellow")

#lines(density(Y))

lines(density(Y), col="red")

```

## ggplot fancy histogram and density curve.

```{r}

ggplot(CO2, aes(x = uptake)) + geom_histogram(aes(x=uptake,y= after_stat(density)),
    binwidth = ) + 
    geom_density(aes(x=uptake,y=after_stat(density)), col = "red")

```

## ggplot just the density curve.

```{r}

ggplot(CO2, aes(x = uptake)) + geom_density(col = "blue", fill = "yellow")

ggplot(CO2, aes(x = uptake)) + geom_density(col = "blue") # no fill colour
```

## ggplot fancy histogram and density curve.

```{r}
ggplot(CO2, aes(x = uptake)) + geom_histogram(aes(x=uptake ,y= after_stat(density)),
  binwidth = 3) + geom_density(aes(x=uptake,y=after_stat(density)), col = "red") +
  ggtitle("Distribution of Carbon Dioxide Uptake") + 
  labs(x = expression(paste("carbon dioxide uptake rates (", mu, "mol/", m^2, " sec)"))) 
```

## Theoretic normal distribution density curve.

```{r}
mu <- 20
sigma <- 5

x <- seq(mu - 3*sigma, mu + 3*sigma, by=0.05)
#x
y <- dnorm(x, mean = mu, sd = sigma)
#y=f(x)
y
normaldata <- as.data.frame(cbind(x, y))

ggplot(normaldata, aes(x = x, y = y)) + geom_line()

```

## Example from lecture note.

```{r}

sampleData=read.csv("variable.X.Sample.csv")

head(sampleData)

X=sampleData$X

#Check if data is normally distributed.

#Fast histogram using basic R graphics.

hist(X, main="Distribution of X", xlab="Values of X", prob=TRUE)
lines(density(X), col="red")

# qq-norm plot

qqnorm(X)
qqline(X)

# A statistical test called Shapiro-Wilk normality test

shapiro.test(X)


# the middle 80%

low <- (1 - .8)/2
low
high <- 1 - low

#quantile(X, 0.10)

#quantile(X, 0.90)

quantile(X, c(0.1, 0.9))


quantile(X, c(0.3, 0.7))



quantile(X, c(0.16, 0.84)) # range for 1 sd from mean, empirical rule

quantile(X, c(0.025, 0.975)) # range for 2 sd from mean, empirical rule

xbar=mean(X)
s=sd(X)
c(xbar-2*s, xbar+2*s)

qnorm(c(0.1, 0.9), xbar, s) #return the value x such that the probability that a  normal random distribution with mean=xbar and s.d.=s is less than or equal to c(0.1, 0.9).
 

pnorm(41.49, mean=xbar, sd=s) #gives the cumulative probability P(X<=x) for a normal distribution with mean=xbar and s.d.=s

pnorm(36.41, xbar, s)

```


Example: average lifetime for lightbulbs

```{r}
# 1. We compute the estimated population mean μ using sample mean x. This will serve as the mean for the approximate normal sampling distribution.
n <- 80
xbar <- 103.4
# 2. We compute the estimated standard error (ese) of the sample mean, which serves as the standard deviation of the approximate normal sampling distribution.
s <- 22.5
ese <- s/sqrt(n)

# 3. We use qnorm() function to find 92% confidence interval for the population mean.

conflevel <- .92
low <- (1 - conflevel)/2
high <- 1 - low

qnorm(c(low, high), mean = xbar, sd = ese)

```


proportion of red cars in Victoria

```{r}
# 1. What is the observed value of the sample statistic/ the estimate of the population proportion?

n <- 320
x <- 62

p_hat <- x/n

p_hat
# 2. Determine a 95% confidence interval for the population proportion p (round to 3 decimal places):

ese <- sqrt(p_hat*(1 - p_hat)/n)
ese

conflevel <- .95

low <- (1 - conflevel)/2
high <- 1 - low

round(qnorm(c(low, high), mean = p_hat, sd = ese), 3)

moe <- (qnorm(high, mean = p_hat, sd = ese) - qnorm(low, mean = p_hat, sd = ese))/2
moe

```

# Determine a 95% confidence interval for the population mean of variable X.

```{r}

sampleData=read.csv("variable.X.Sample.csv")
head(sampleData)

X=sampleData$X
x_bar = mean(X)
s = sd(X)
n= length(X)

ese = s/sqrt(n) 

conflevel <- .95
low <- (1 - conflevel)/2
high <- 1 - low

qnorm(c(low, high), mean = x_bar, sd = ese)

#estimate is x_bar
#margin of error calculation

moe <- (qnorm(high, mean = x_bar, sd = ese) - qnorm(low, mean = x_bar, sd = ese))/2
moe

#double check

x_bar - moe
x_bar + moe

```

```{r}
x<-seq(-5,5,.01)
y.norm0<-dnorm(x,mean=0,sd=1)
y.norm1<-dnorm(x,mean=1,sd=1)
y.norm2<-dnorm(x,mean=2,sd=1)


# set up the plot area
plot(c(min(x),max(x)),c(min(y.norm0,y.norm1,y.norm2),max(y.norm0,y.norm1,y.norm2)),type="n",
xlab="x",ylab="Density function: f(x)")
title("The Normal distribution (sd=1) with different means")
lines(x,y.norm0, col="red")
lines(x,y.norm1,col="purple")
lines(x,y.norm2,col="orange")
legend(x=c(-5.1,-3.1),y=c(.4,.28),legend=c('N(0,1)','N(1,1)','N(2,1)'),fill=c('red','purple','orange'))
```

```{r}
x<-seq(-5,5,.01)
y.norm0<-dnorm(x,mean=0,sd=1)
y.norm1<-dnorm(x,mean=0,sd=1.5)
y.norm2<-dnorm(x,mean=0,sd=2)


# set up the plot area
plot(c(min(x),max(x)),c(min(y.norm0,y.norm1,y.norm2),max(y.norm0,y.norm1,y.norm2)),type="n",
xlab="x",ylab="Density function: f(x)")
title("The Normal distribution (mean=0) with different standard deviation")
lines(x,y.norm0)
lines(x,y.norm2,col="green")
lines(x,y.norm1,col="blue")
legend(x=c(2,4.2),y=c(.4,.28),legend=c('N(0,1)','N(0,1.5)','N(0,2)'),fill=c('black','blue','green'))
```

```{r}
s = sd(X)
n= length(X)
ese = s/sqrt(n) 
ese
```

```{r}

```

