---
title: "Untitled"
format: docx
editor: visual
---

## Grades raw

```{r}
students <- read.csv(file = "rawgrades.csv", header=TRUE)
dim(students)
head(students)
class(students)
```

```{r}
students <- read.csv(file = "~/UVIC/Stat123/rawgrades.csv", header = TRUE)
students
```

```{r}
install.packages("readxl")
```

```{r}
table(students[,2])
```

```{r}
gradetable <- table(students[,2])
barplot(gradetable, main="grade dist", xlab="letter grade")
library(readxl)
```

```{r}

```

A factor has categorical data while a vector doesn't need to.

A factor can also only store one data type at a time

```{r}
students$Grade <- factor(students$Grade, levels <- c("A+", "A", "A-", "B+", "B", "B-", "C", "D", "F"))
levels(students$Grade)
table(students[,2])
```

**nominal = unordered classifying**

**ordinal = ordered classifying**

```{r}
pie(gradetable, main = ) #etc. finish this code for pie chart
```

**it is hard for us to plot data in a pie graph/bar graph when the given data is very spread out and non categorical. how we can plot these is by creating categories of ranges that contain these numbers.**

$$
The\ area\ of\ a\ circle\ is\ A = \pi r^2.
$$

$$
x = \frac{-b \pm \sqrt{b^2-4ac}}{2a}
$$

$$
\sum_{k=1}^n a_k
$$

```{r}

```
