---
title: "PracticeQuestionsMT2"
format: docx
editor: visual
---

## questions

1\) find the median of the following vectors c(83, 83, 63, 2000, 74, 83, 79, 61, 82, 85, 62), c(83, 12, 3, 1, 99, 101). hint, do not use a built in function like median(). then do it with median()

```{r}
v1 <- c(83, 83, 63, 2000, 74, 83, 79, 61, 82, 85, 62)
v1 <- sort(v1, decreasing = TRUE)
v1[length(v1)/2 +1]
v2 <- c(83, 12, 3, 1, 99, 101)
v2 <- sort(v2, decreasing = FALSE)

v2 <- (v2[length(v2)/2] + v2[length(v2)/2 + 1])/2
v2
median(v2)
```

2\) create a dataframe of the following vectors: score \<- c(82, 60, 77, 90, 70)wt1 \<- c(.15, .10, .15, .10, .50)wt2 \<- c(30, 50, 40, 25, 80). then from the data frame find the mean of one of the columns, do this using cbind and rbind

```{r}
score <- c(82, 60, 77, 90, 70)
wt1 <- c(.15, .10, .15, .10, .50)
wt2 <- c(30,50,40,25,80)
df <- as.data.frame(cbind(score, wt1, wt2))
df2 <- as.data.frame(rbind(score,wt1,wt2))
mean(df$score)
mean(df$score*df$wt1)
(sum(df$score) + sum(df$wt1))/(length(df$score) + length(df$wt1))
```

3\) for this vector: c(2, 4, 8), find the geometric mean in 2 ways.

```{r}
vec <- c(2,4,8)
geom <- (prod(vec))**(1/length(vec))
geom2 <- exp((1/length(vec))*log(prod(vec)))
geom
geom2
```

4\) When should you use geometric, harmonic, or typical mean?:

you should use the geometric mean for things like SKEWED DISTRIBUTIONS, population increase, ratios. Use harmonic for things like speed and distance sometimes time periods too.

5\) use geometric mean on the stock.csv data set (hint remove na and use -1) on stock\$ratio

```{r}

```

6\) find the mean of 3 columns of the stocks.csv data set (hint remove na)

```{r}

```

7\) read the speed.csv data set and then apply the harmonic mean on the time and distance

```{r}

```

8\) find the min and max (range) of: c(83, 83, 63, 2000, 74, 83, 79, 61, 82, 85, 62) using min() and max(), and another function

```{r}

```

9\) use cumsum on the following to find the different percentiles for each item: c(81, 93, 63, 2000, 74, 83, 79, 61, 82, 85, 62)

```{r}

```

10\) load the table of the previous vector into a data frame, and add columns representing cumulative frequence and relative frequency. understand what they mean

```{r}

```

11\) use min, max, median, IQR, quantile, summary on the bulit in airquality dataset. (try putting numbers and vectors with the quantile function

```{r}

```

12\) use vanilla R to create a box blot (box and whisker plot) for temp and month in the air quality data set

```{r}

```

13\) do the same but with ggplot

```{r}

```

14\) find the mean deviation, absolute mean deviation, and the mean square deviation

```{r}

```

15\) find the standard deviation and variance of temp in 2 different ways.

```{r}

```

16\) use LETTERS to make a data frame of letters and reps.
