---
title: "Lab 1"
author: "Evan O'Toole"
date: "January, 24th, 2025"
output:
  html_document: default
  pdf_document: default
---

# Likelihood methods for the Binomial distribution

* Put your name in the author section.
* Execute each chunk of code to ensure that your code works properly.  The TA will review the code
with you.
* Sometimes one of your chucks of code will not compile properly, but you must hand
your document in.  In that case, 'Comment' out the R code that is not working properly using #.
Do that now.  For the chuck of code in Section number 4, below, put #  at the start of each line of R 
code. 
* Compute question 6 below.  Execute the question 6 code to ensure that it works properly.  If you cannot
compile it without errors before the end of the class, comment out the chunk.
* Then HTML knit the entire document.
* Open the HTML document in a browser.
* Save the HTML file as a pdf file and upload it to the Practice Lab Assignment Activity in the 
Lab section of CourseSpaces. 


## 1.  First we generate 1 observation from the Binomial(100, theta=.07) distribution


```{r data}
#This is a comment inside of a chunk of R code.

set.seed(54321)   #setting the random number generator seed SHOULD always be done before simulating
# Ask your Lab instructor to explain set.seed()
bdat <- rbinom(1, size=100, prob=.07 )   # see ?rbinom for information
bdat
```

## 2.  We know that the MLE for the probability of success, theta, is the sample proportion.

```{r MLE_theta}
thetahat <- bdat/100
thetahat
```

The MLE of theta is `r thetahat`.  

## 3.  Compute the vector of Log-Likelihoods for a sequence of values for theta given the one data point bdat.


```{r Log-Like}
theta <- seq(0.02, 0.14, by=.005)   #creates a vector of values from .02 to .14 in steps of .005
bloglike <- dbinom(bdat, size=100, prob=theta, log=TRUE)  #computes log of binomial probabilities
cbind(theta, bloglike)
```

## 4. Plot the Log-Likelihood function versus theta.

```{r plot}
plot(bloglike~theta, ylab='Log-Likelihood', xlab='theta', type='b')
title('Binomial Log-likelihood for Lab 2')
```

## 5.  Generate n=10 observations from the same distribution and plot the log-likelihood function given the n observations.

```{r Sample}
set.seed(54321)
n <- 10
bdat2 <- rbinom(n, size=100, prob=.07 )   # this generates n binomial observations
bdat2
theta <- seq(0.02, 0.14, by=.005)

bloglike2 <- 0
for (i in 1:n){
bloglike2 <- bloglike2 + dbinom(bdat2[i], size=100, prob=theta, log=TRUE)  
      #computes the sum of the log of binomial probabilities over the observations
} 

cbind(theta, bloglike2)
plot(bloglike2~theta, ylab='Log-Likelihood', xlab='theta', type='b')
title(paste('Binomial Log-likelihood for Lab 2, n=', n, sep=''))



```

## 6. Now it is your turn.  In the R code chunk below, redo question 5. for n=100 observations from the same distribution.  How does the log-liklihood function compare with that in question 5?  Write your comments in this section.

Comments:  

```{r YourTurn}
set.seed(54321)
n <- 100
bdat2 <- rbinom(n, size=100, prob=.07 )   # this generates n binomial observations
bdat2
theta <- seq(0.02, 0.14, by=.005)

bloglike2 <- 0
for (i in 1:n){
bloglike2 <- bloglike2 + dbinom(bdat2[i], size=100, prob=theta, log=TRUE)  
      #computes the sum of the log of binomial probabilities over the observations
} 

cbind(theta, bloglike2)
plot(bloglike2~theta, ylab='Log-Likelihood', xlab='theta', type='b')
title(paste('Binomial Log-likelihood for Lab 2, n=', n, sep=''))
```





